Toronto Homicides and Shootings (To be Changed)

### 2024-09-25 paper/paper.qmd
Updated graphs and tables in paper

# Results {#sec-results}
## Number of Homicide and Shooting Cases in Toronto from 2018 to 2023 in Toronto
```{r}
#| include: false
#| warning: false
#| message: false
# Add up total number of cases per neighbourhood
crime_per_neighbourhood <- 
  cleaned_crime_data |>
  # Calculate the total sum of cases across all neighbourhoods 
  mutate(
    num_cases_total = sum(num_of_cases)
  ) |>
  select(neighbourhood, homicide_or_shooting, num_of_cases, num_cases_total) |>
  # Groups data by type of crime, and neighbourhood
  group_by(neighbourhood, homicide_or_shooting) |>
  # Create new columns for number of cases for homicides and shootings per neighbourhood. Also calculate the proportion as well. 
  mutate (
    num_cases_total_per_neighbourhood_per_crime = sum(num_of_cases),
    num_cases_total_per_neighbourhood_per_crime_proportion = round(num_cases_total_per_neighbourhood_per_crime / num_cases_total, 3)
  ) |>
  ungroup() |>
  group_by(neighbourhood) |> 
  mutate(
    num_cases_total_per_neighbourhood = sum(num_of_cases),
    num_cases_total_per_neighbourhood_proportion = round(num_cases_total_per_neighbourhood / num_cases_total, 3)
  ) |>
  # Obtains unique rows
  unique() |>
  # Arrange based on the total number of cases per neighbourhood
  arrange(desc(num_cases_total_per_neighbourhood)) |>
  tibble() |>
  select(neighbourhood, homicide_or_shooting, num_of_cases, num_cases_total, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_proportion, num_cases_total_per_neighbourhood_per_crime_proportion)

@@

```{r}
#| label: tbl-totalcasestable
#| fig-cap:  Number of Homicide and Shooting Cases in Toronto from 2018 to 2023
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: tbl-highestcasesshootingstable
#| fig-cap:  Six Toronto Neighbourhoods with the Highest Cases of Shootings from 2018 to 2023
#| echo: false
#| warning: false
#| message: false

# Creates table
crime_per_neighbourhood |>
  filter(homicide_or_shooting == "shooting") |>
  arrange(desc(num_cases_total_per_neighbourhood_per_crime)) |>
  select(neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_per_crime_proportion) |>
  unique() |>
  rename(
    "Total number of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime,
    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime_proportion,
    "Neighbourhood" = neighbourhood
  ) |>
  head(6) |>
  kable()
```

```{r}
#| label: tbl-highestcaseshomicidestable
#| fig-cap:  Six Toronto Neighbourhoods with the Highest Cases of Homicides from 2018 to 2023
#| echo: false
#| warning: false
#| message: false

# Creates table
crime_per_neighbourhood |>
  filter(homicide_or_shooting == "homicide") |>
  arrange(desc(num_cases_total_per_neighbourhood_per_crime)) |>
  select(neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_per_crime_proportion) |>
  unique() |>
  rename(
    "Total number of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime,
    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime_proportion,
    "Neighbourhood" = neighbourhood
  ) |>
  head(6) |>
  kable()
```

```{r}
#| label: tbl-highestcasestable
#| fig-cap:  Six Toronto Neighbourhoods with the Highest Cases of Homicides and Shootings Combined from 2018 to 2023
#| echo: false
#| warning: false
#| message: false

# Creates table
crime_per_neighbourhood |>
  filter(homicide_or_shooting == "homicide") |>
  arrange(desc(num_cases_total_per_neighbourhood)) |>
  select(neighbourhood, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_proportion) |>
  unique() |>
  rename(
    "Total number of cases (2018 to 2023)" = num_cases_total_per_neighbourhood,
    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_proportion,
    "Neighbourhood" = neighbourhood
  ) |>
  head(6) |>
  kable()
```

@@

crime_per_neighbourhood |> 
  select(neighbourhood, homicide_or_shooting, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime) |>
  arrange(desc(num_cases_total_per_neighbourhood)) |>
  select(-c(num_cases_total_per_neighbourhood)) |>
  unique() |>
  head(12) |>
  ggplot(mapping = aes(x = neighbourhood, y = num_cases_total_per_neighbourhood_per_crime, fill = homicide_or_shooting)) +
  geom_bar(stat="identity", position = "dodge2") +
  labs(x = "Year", y = "Number of Cases", fill = "Type of Crime") +
  theme_minimal() + 
  # Reference https://stackoverflow.com/questions/1330989/rotating-and-spacing-axis-labels-in-ggplot2 to figure out how to rotate text along x axis.
  theme(legend.position = "bottom", plot.title = element_text(size = 9, face = "bold"), legend.title = element_text(size = 9), legend.text = element_text(size = 7), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))  + scale_fill_viridis_d()
```

@@

```{r}
#| label: fig-shootings-map
#| fig-cap: Location of police facilities in relation to Toronto neighbourhoods with the highest number of shootings from 2018 to 2023.
#| echo: false
#| warning: false
#| message: false
# Reference the following for code:
# https://cengel.github.io/R-spatial/mapping.html
ggplot() + 
    # To help convert between GeoJSON to sf  https://cran.r-project.org/web/packages/geojsonsf/vignettes/geojson-sf-conversions.html
  # Fill in area regarding the number of cases 
  geom_sf(data = cleaned_crime_data |> filter(homicide_or_shooting == "shooting"), aes(fill = num_of_cases, geometry = geometry)) +
  # Changes the font of the title and removes the x and y axis 
  # Reference to figured out how to add labels to maps https://tmieno2.github.io/R-as-GIS-for-Economists/create-maps.html
  theme_minimal() + 
  labs(
    x = "Longitude",
    y = "Latitude",
    title = "Number of shootings across Toronto neighbourhoods (2018 to 2023)",
    fill = "Number of cases"
  ) +
  geom_sf_text(
    data = cleaned_crime_data |> filter(num_of_cases >= 14),
    aes(label = hood_id),
    check_overlap = TRUE,
    size = 2, 
    color = "white"
    ) +
    theme(plot.title = element_text(size = 9, face = 'bold'), 
          plot.subtitle = element_text(size = 9),
          rect = element_blank(), 
          axis.text.x = element_blank(), 
          axis.text.y = element_blank(), 
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(), 
          axis.ticks = element_blank(), 
          legend.title = element_text(size = 8),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()
          ) +
  # Add in police locations
  geom_point(data = cleaned_police_location,
    aes(geometry = geometry, color = "Police Facility"), stat = "sf_coordinates"
  ) + labs(color = "Location")

>>>>

# Results {#sec-results}
## Number of Homicide and Shooting Cases in Toronto from 2018 to 2023 in Toronto
```{r}
#| include: false
#| warning: false
#| message: false
# Add up total number of cases per neighbourhood
crime_per_neighbourhood <- 
  cleaned_crime_data |>
  # Calculate the total sum of cases across all neighbourhoods 
  mutate(
    num_cases_total = sum(num_of_cases)
  ) |>
  select(neighbourhood, hood_id, homicide_or_shooting, num_of_cases, num_cases_total) |>
  # Groups data by type of crime, and neighbourhood
  group_by(neighbourhood, homicide_or_shooting) |>
  # Create new columns for number of cases for homicides and shootings per neighbourhood. Also calculate the proportion as well. 
  mutate (
    num_cases_total_per_neighbourhood_per_crime = sum(num_of_cases),
    num_cases_total_per_neighbourhood_per_crime_proportion = round(num_cases_total_per_neighbourhood_per_crime / num_cases_total, 3)
  ) |>
  ungroup() |>
  group_by(neighbourhood) |> 
  mutate(
    num_cases_total_per_neighbourhood = sum(num_of_cases),
    num_cases_total_per_neighbourhood_proportion = round(num_cases_total_per_neighbourhood / num_cases_total, 3)
  ) |>
  # Obtains unique rows
  unique() |>
  # Arrange based on the total number of cases per neighbourhood
  arrange(desc(num_cases_total_per_neighbourhood)) |>
  tibble() |>
  select(neighbourhood, hood_id, homicide_or_shooting, num_of_cases, num_cases_total, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_proportion, num_cases_total_per_neighbourhood_per_crime_proportion)

@@

```{r}
#| label: tbl-totalcasestable
#| tbl-cap: Number of Homicide and Shooting Cases in Toronto from 2018 to 2023
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: tbl-highestcasesshootingstable
#| tbl-cap:  Six Toronto Neighbourhoods with the Highest Cases of Shootings from 2018 to 2023
#| echo: false
#| warning: false
#| message: false

# Creates table
crime_per_neighbourhood |>
  filter(homicide_or_shooting == "shooting") |>
  arrange(desc(num_cases_total_per_neighbourhood_per_crime)) |>
  select(neighbourhood, hood_id, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_per_crime_proportion) |>
  unique() |>
  rename(
    "Total number of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime,
    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime_proportion,
    "Neighbourhood ID" = hood_id,
    "Neighbourhood" = neighbourhood
  ) |>
  head(6) |>
  kable()
```

```{r}
#| label: tbl-highestcaseshomicidestable
#| tbl-cap:  Six Toronto Neighbourhoods with the Highest Cases of Homicides from 2018 to 2023
#| echo: false
#| warning: false
#| message: false

# Creates table
crime_per_neighbourhood |>
  filter(homicide_or_shooting == "homicide") |>
  arrange(desc(num_cases_total_per_neighbourhood_per_crime)) |>
  select(neighbourhood, hood_id, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_per_crime_proportion) |>
  unique() |>
  rename(
    "Total number of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime,
    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime_proportion,
    "Neighbourhood ID" = hood_id, 
    "Neighbourhood" = neighbourhood
  ) |>
  head(6) |>
  kable()
```

```{r}
#| label: tbl-highestcasestable
#| tbl-cap:  Six Toronto Neighbourhoods with the Highest Cases of Homicides and Shootings Combined from 2018 to 2023
#| echo: false
#| warning: false
#| message: false

# Creates table
crime_per_neighbourhood |>
  filter(homicide_or_shooting == "homicide") |>
  arrange(desc(num_cases_total_per_neighbourhood)) |>
  select(neighbourhood, hood_id, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_proportion) |>
  unique() |>
  rename(
    "Total number of cases (2018 to 2023)" = num_cases_total_per_neighbourhood,
    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_proportion,
    "Neighbourhood ID" = hood_id,
    "Neighbourhood" = neighbourhood
  ) |>
  head(6) |>
  kable()
```

@@

crime_per_neighbourhood |> 
  select(hood_id, homicide_or_shooting, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime) |>
  arrange(desc(num_cases_total_per_neighbourhood)) |>
  select(-c(num_cases_total_per_neighbourhood)) |>
  unique() |>
  head(12) |>
  ggplot(mapping = aes(x = as.character(hood_id), y = num_cases_total_per_neighbourhood_per_crime, fill = homicide_or_shooting)) +
  geom_bar(stat="identity", position = "dodge2") +
  labs(x = "Neighbourhood ID", y = "Number of Cases", fill = "Type of Crime") +
  theme_minimal() + 
  # Reference https://stackoverflow.com/questions/1330989/rotating-and-spacing-axis-labels-in-ggplot2 to figure out how to rotate text along x axis.
  theme(legend.position = "bottom", plot.title = element_text(size = 9, face = "bold"), legend.title = element_text(size = 9), legend.text = element_text(size = 7))  + scale_fill_viridis_d()
```

@@

```{r}
#| label: fig-shootings-map
#| fig-cap: Location of police facilities in relation to six Toronto neighbourhoods with the highest number of shootings from 2018 to 2023.
#| echo: false
#| warning: false
#| message: false
# Reference the following for code:
# https://cengel.github.io/R-spatial/mapping.html
ggplot() + 
    # To help convert between GeoJSON to sf  https://cran.r-project.org/web/packages/geojsonsf/vignettes/geojson-sf-conversions.html
  # Fill in area regarding the number of cases 
  geom_sf(data = cleaned_crime_data |> filter(homicide_or_shooting == "shooting"), aes(fill = num_of_cases, geometry = geometry)) +
  # Changes the font of the title and removes the x and y axis 
  # Reference to figured out how to add labels to maps https://tmieno2.github.io/R-as-GIS-for-Economists/create-maps.html
  theme_minimal() + 
  labs(
    x = "Longitude",
    y = "Latitude",
    title = "Number of shootings across Toronto neighbourhoods (2018 to 2023)",
    fill = "Number of cases"
  ) +
  geom_sf_text(
    data = cleaned_crime_data |> filter(num_of_cases >= 14),
    aes(label = hood_id),
    check_overlap = TRUE,
    size = 2, 
    color = "white"
    ) +
    theme(plot.title = element_text(size = 9, face = 'bold'), 
          plot.subtitle = element_text(size = 9),
          rect = element_blank(), 
          axis.text.x = element_blank(), 
          axis.text.y = element_blank(), 
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(), 
          axis.ticks = element_blank(), 
          legend.title = element_text(size = 8),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()
          ) +
  # Add in police locations
  geom_point(data = cleaned_police_location,
    aes(geometry = geometry, color = "Police Facility"), stat = "sf_coordinates"
  ) + labs(color = "Location")
